---
title: "Final Project - Explring Starbucks database"
author: "Kovacs_Adel"
date: "1/1/2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Kovacs Adel: Final Project
## PSZM21-MO-KUT-104:2	Komplex adatelemzési eljárások-Adatelemzés R-programnyelven

I have choosen the Starbucks dataset from TidyTuesday
I like coffee and I know some basic things about it (like it has caffeine, you can add milk, you can have a more concentrated espresso or a "long one", an Americano), but I have seen some data that I am not very familiar with. 
The goal is to see what variables predict some attributes for different Sturbucks beverages. 

### Packages
I have loaded the necessary packages. I have came back here every time I needed an other one

```{r, message=FALSE}
library(tidyverse)
library(corrplot)
library(performance)
library(ggpubr)
library(broom)
library(car)
library(multcompView)
```

### Loading the data

```{r}
starbucks <- read.csv(url("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-12-21/starbucks.csv"))
View(starbucks)
```

### Checking and modifying variables

I checked the type of variables I have, I am going to change them if it seems necessary

```{r}
str(starbucks)
```

Checking for any coding mistakes or extraordinary data mainly with graphs (pairing the variables and checking two at a time - also good for exploring any possible connection or correlation). Wherever I find an extreme outlier I check the variable type and will filter out if it does not make sense

```{r, message=FALSE}
ggplot(starbucks, aes(x = serv_size_m_l, y = calories)) +
  geom_point () +
  geom_smooth(method = "lm", se = FALSE)

ggplot(starbucks, aes(x = total_fat_g, y = saturated_fat_g)) +
  geom_point()

ggplot(starbucks, aes(x = trans_fat_g, y = cholesterol_mg)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

ggplot(starbucks, aes(x = sodium_mg, y = total_carbs_g)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

ggplot(starbucks, aes(x = fiber_g, y = sugar_g)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

ggplot(starbucks, aes(caffeine_mg, calories)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

### Decisions about the coding errors and outliers

Service size equals zero does not seem to be right, I filtered them out. 

Zero fat seems to be possible, I have not changed those. 

There is an item with extreme high trans fat (2mg), I filtered it out. I do not really know if it is possible, but can be a quite big effect on a fitted model, so I have decided to extract it. 

Other variables and values seemed to be right.

```{r}
starbucks <- starbucks %>%
  filter(serv_size_m_l > 0, trans_fat_g < 2)
```

I would like the *whip* to be a factor and rename the levels (0 becomes no, 1 becomes yes). Also checking what I did during the process

```{r}
starbucks <- 
  starbucks  %>%
mutate(whip = as.factor(whip))

is.factor(starbucks$whip)

starbucks <- starbucks %>%
  mutate(whip = fct_recode(whip, 'no' = '0', 'yes' ='1'))

levels(starbucks$whip)
```

I would also like *size* to be a factor, because it has definite levels and it is better as factor then as character to work with

```{r}
starbucks <- starbucks %>%
  mutate(size = as.factor(size))

levels(starbucks$size)
```

### Building models

I saw some not so flat lines when I made the plots, so I am checking the correlations between all the numeric variables (for this I need to filter out the two factors and the character name)

```{r}
summary(starbucks)
starbucks_cor <- starbucks %>%
  select(3, 5:14)
summary(starbucks_cor)
str(starbucks_cor)

correlations <- starbucks_cor %>%
  cor()
```

I am visulazing the correlations to see it better

```{r}
corrplot(correlations, method = "circle")
```
### Understanding the correlation matrix

The diagonal is every variable with itself, they are a perfect +1 correlation as they should be 
As it was predictable different measurments of fat (*trans_fat_g*, *total_fat_g*, *saturated_fat_g*) have a strong correlation (r = 0.77 - 0.96)
Carbs (*total_carbs_g*) and sugar (*sugar_g*) also have a predicted correlation (r = 0.99), the almost perfect correaltion makes sense, since the coffees containing coffee, milk, whip and sugar the only and main source of carbs is sugar. 
Sugar and fat has a connection with calories, moderate between trans fat and calories (r = 0.66), and strong with the others (r = 0.76 - 0.89)
Just by taking a glance at the matrix, it seems that calories have the most connection to the other variables, so later, I will try to fit a model to predict calories in coffees. 

### Checking different size and whipped cream in relation to calories

As I have decided to work on finding out what makes a coffee with more calories, I would also like to see if it is true that whiped creamed beverages and bigger beverages have more calories

I am using independent sample t-test to see if coffees with shipped cream are more heavy in calories thank the ones without
Nullhypothesis is that there is no difference, alternative hypothesis is that the whipped creamed have more calories

I am checking the normality

```{r}
ggdensity(starbucks$calories, 
          main = "Density plot of calories",
          xlab = "Calories of coffees")
ggplot(starbucks, aes(x = calories, fill = whip)) +
  geom_histogram()
```

It seems skewed so I am running a Shapiro-Wilk test

```{r}
  shapiro.test(starbucks$calories)
```

The Shapiro-Wilk came back significant so I need to transfer my data

I am calculating the model

```{r}
whip_t_test <- wilcox.test(calories ~ whip, data = starbucks)
whip_t_test
```

It seems like we have a significant difference in the calories of whipped and not whipped coffees. Z = 27426, p < 0.001

I checked a boxplot to see graphically which coffees have more calories (although I can see the means above)

```{r}
plot(calories ~ whip, data = starbucks)
```

It is the whipped, so we can state that coffees with whipped cream contain more calories (mean = 371.22 vs. mean of the no whipped cream coffees = 188.54)

Now I take a look at the size of a coffee related to its calories
I am bulding a one-way ANOVA model to find out if there are any differences between the groups, and first I check the assumptions
My nullhypothesis is that there are no differences
My alternative hypothesis is that there are differences

Checking for outliers graphically

```{r}
ggboxplot(starbucks, x = "size", y = "calories",
          ylab = "Calories", xlab = "Size")

outlier <- starbucks %>%
  filter (size == "trenta", calories < 10)
outlier
```

I have found an outlier in size "trenta" which turned out to be a cold brewed coffee and it makes sense, since as far as I know it does not contain any milk or sugar, so I left it in the data

I am checking the homogenity of variances as an assumption
I would like them to be not significantly different (I have checked both the Levene's and the Bartlett)
```{r}
leveneTest(calories ~size, starbucks)

bartlett.test(calories ~size, starbucks)
```

Unfortunately they are significantly different
I need to use non parametric or robust ways of ANOVA (I found a helpful article [here](https://dornsife.usc.edu/assets/sites/239/docs/WRS2.pdf))

```{r}
starbucks_anova <- kruskal.test(calories ~ size, data = starbucks)
starbucks_anova
```

I found a significant model (H(4) = 273.35 p < 0.001)

To find out which groups differ from each other I have run a robust pairwise t-test with Bonferroni correction
It threw me an error saying _'x' must be numeric_ so I tried to change that, but was not successful (fct_recode error: _`f` must be a factor (or character vector)_)
(To this I needed some research to see which size is bigger than the other)

```{r}
# pairwise.wilcox.test(starbucks$size, g = starbucks$calories)

#starbucks %>%
#  fct_recode(size, short = "1", tall = "2", grande = "3", venti = "4", trenta = "5" ) %>%
#  mutate(size = as.numeric(size))

is.factor(starbucks$size)
```

I have produced a boxplot to visualize it

```{r}
boxplot(calories ~ size, data = starbucks)
```

### Fitting a complex model to predict calories in a coffee

Lastly I wanted to fit a complex model to predict calories in a coffee 
I contained every variable that seemed to be correlated to calories

First I built a less complex model (with obvious elements like sugar and whip) and checked the assumptions

```{r}
model1 <- lm(calories ~ whip + sugar_g + total_fat_g, starbucks)
summary(model1)

check_model(model1)

model1 %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  head()
```

Assumptions are not violated and I have a significant model where sugar, whipcream and total fat predicts the calorie of the drink (F(1111,3) = 1.136e+04, p < 0.001) and all three of the predictors are significant elements (p < 0.001)

Next I built a model with total carbs to compare them. 

```{r}
model2 <- lm(calories ~ whip + sugar_g + total_fat_g + total_carbs_g, starbucks)
summary(model2)

check_model

model2 %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  head()
```

The model would be significant, but the collinearity assumption is violated: sugar and total carbs are correlated too highly, so I have to drop one of them and I am back to model 1

I am looking an other model with caffeine in it, I think if I compare model 1 with model 3 with caffeine, then model1 should be better

```{r}
model3 <- lm(calories ~ whip + sugar_g + total_fat_g + caffeine_mg, starbucks)
summary(model3)

check_model(model3)

model3 %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  head()
```

The assumptions are met and I got a significant model (F(1110,4)=8704, p < 0.001), where all of the predictors are significant, even the caffeine (p < 0.001)

I am comparing the models without and with caffeine

```{r}
compare <- anova(model1, model3)
compare
```

Turns out that model3 with caffeine is a significantly better one that the model1 without it (F(1) = 24.51, p < 0.001)
I assume that bigger sizes have more caffeine but also more milk and sugar

### A plot about the calories of different sized and whipped coffees

I wanted to see the relation between size, whipped cream and calories of coffees

```{r}
starbucks_to_plot <- starbucks %>%
  group_by(size) %>%
  count(mean(calories))

ggplot(starbucks_to_plot, aes(x = reorder(size, -n), y = n)) +
  geom_col()  +
  scale_colour_brewer()  +
  theme_classic() +
  labs (title = "Mean calories of Starbucks coffees by size", x = "Size", y = "Calories") +
  theme (legend.position ="bottom")
```

```{r}

```